---
title: "Finding the Number of CPGs"
format:
  html:
    toc: true
    toc-depth: 5
    html-math-method: katex
    embed-resources: true
    smooth-scroll: true
editor: visual
execute:
  echo: true
  include: true
---

# Set up

```{r set-up}
library(IlluminaHumanMethylation450kanno.ilmn12.hg19)
library(rio)
library(tidyverse)
library(ggVennDiagram)
library(gridExtra)
library(RColorBrewer)
library(janitor)
library(GenomicRanges)
```

# Gene Lists

Here we read in pathway gene lists and see how many genes are in each pathway.

```{r}
kegg_immune <- import("data/pathway_gene_lists/human_KEGG_immune_pathways_unique_gene_list.csv")
cat("KEGG immune pathway has ", nrow(kegg_immune), "genes\n")

kegg_cvd <- import("data/pathway_gene_lists/KEGG_CVD_unique_gene_symbols.csv")
cat("KEGG CVD pathway has ", nrow(kegg_cvd), "genes\n")

kegg_endocrine <- import("data/pathway_gene_lists/KEGG_endocrine_metabolic_unique_gene_symbols.csv")
cat("KEGG endocrine pathway has ", nrow(kegg_endocrine), "genes\n")

# kegg_immune_cvd_endocrine_genes<- c(kegg_immune$gene_symbol, kegg_cvd$gene_symbol, kegg_endocrine$gene_symbol)

atherosclerosis_df <- import("data/pathway_gene_lists/hsa05417_Lipid_and_atherosclerosis_genes.csv")

# We decided to look at just atherosclerosis genes
unique_kegg_ids <- unique(atherosclerosis_df$gene_symbol)
cat("atherosclerosis pathway has ", length(unique_kegg_ids), "genes\n")
```

# Illumina EPIC 450k

Here we read the EPIC 450k manifest file.

```{r}
anno_450k <- getAnnotation(IlluminaHumanMethylation450kanno.ilmn12.hg19)
anno_450k_df <- as.data.frame(anno_450k)

anno_450k_subset <- anno_450k_df %>%
  select(chr, pos, Name, UCSC_RefGene_Name)
cat("450k Array has ", nrow(anno_450k_subset), " rows\n") #485512      4

anno_450k_subset_long <- anno_450k_subset %>%
    separate_rows(UCSC_RefGene_Name, sep = ";") %>%
    distinct()
cat("450k Array has ", nrow(anno_450k_subset_long), " rows of genes") #485512      4
```

# Common Genes in 450k Manifest

Of the 216 genes in our pathway, we have 202 in the 450k array manifest

```{r}
anno_genes <- anno_450k_subset_long %>%
  filter(!is.na(UCSC_RefGene_Name), UCSC_RefGene_Name != "") %>%
  pull(UCSC_RefGene_Name) %>%
  unique()

kegg_genes <- unique_kegg_ids %>%
  na.omit() %>%
  unique()

venn_list <- list(
  `450k` = unique(anno_genes),
  `kegg` = unique(kegg_genes)
)

ggVennDiagram(venn_list)
```

# Number Probes and Genes

```{r}
anno_450k_subset_long_filt <- anno_450k_subset_long %>%
  filter(UCSC_RefGene_Name %in% unique_kegg_ids)  
cat("We filter 450k table for genes in our pathway of interest, ", nrow(anno_450k_subset_long_filt), "rows\n")
# unique(anno_450k_subset_long_filt$chr) 

anno_450k_subset_long_filt_xy <- anno_450k_subset_long_filt%>%
  filter(!chr %in% c("chrY", "chrX"))
cat("We filter 450k table further, removing probes on chromosome X and Y, ", nrow(anno_450k_subset_long_filt_xy), "rows\n")
# unique(anno_450k_subset_long_filt_xy$chr) # chrX/Y removed

cat("So for our genes of interest in 450k manifest table, we have ", n_distinct(anno_450k_subset_long_filt_xy$Name), " distinct probes after filtering out chr XY probes\n")
cat("So for our genes of interest in 450k manifest table, we have ", n_distinct(anno_450k_subset_long_filt_xy$UCSC_RefGene_Name), " distinct genes after filtering out chr XY probes\n")
```

# Our Data

Loading M matrix data for Healthy Start

```{r}
hs_data_Ms <- import("~/Documents/GitHub/healthy_start/data/Borengasser_170160-110104/CB_m_norm_batch_plate.csv")

cat("We have ", length(unique(anno_450k_subset_long_filt_xy$Name)), " probes for the genes of interest in the 450k manifest\n" )
cat("Of those ", length(intersect(names(hs_data_Ms), anno_450k_subset_long_filt_xy$Name)), " probes for genes of interest are also found in our healthy start data\n" )

# Here we filter out health start data for the probes linked to genes of interest
healthy_start_Ms_filtered <- hs_data_Ms %>%
  select(pid, any_of(anno_450k_subset_long_filt_xy$Name))

cat("Dimension of our healthy start data filtered to probes for genes of interest ", dim(healthy_start_Ms_filtered))
saveRDS(healthy_start_Ms_filtered, "healthy_start_Ms_filtered_to_probes_for_genes_of_interest.rds")


```

# Common Genes in Data

Of the 216 genes, we have 194 in our healthy start data.

```{r}
anno_genes2 <- anno_450k_subset_long %>%
  filter(!is.na(UCSC_RefGene_Name), UCSC_RefGene_Name != "") %>%
  filter(Name %in% names(healthy_start_Ms_filtered)) %>%
  pull(UCSC_RefGene_Name) %>%
  unique()

kegg_genes <- unique_kegg_ids %>%
  na.omit() %>%
  unique()

venn_list <- list(
  `our_genes` = anno_genes2,
  `kegg` = kegg_genes
)

ggVennDiagram(venn_list)
```
