---
title: "Healthy Start Power Analysis Using WebPower (R)"
date: today
format:
  docx:
    toc: true
    number-sections: true
    syntax-highlighting: github
  pdf: 
    documentclass: article
  html:
    toc: true
    toc-depth: 5
    html-math-method: katex
    embed-resources: true
    smooth-scroll: true
editor: visual
execute:
  echo: true
  include: true
---

```{r, include = FALSE}
library(WebPower)
```

# WebPower

https://webpower.psychstat.org/models/reg01/

```{r, echo=FALSE}
cat("version:", as.character(packageVersion("WebPower")))
```

```{r, include = FALSE}
# Helper function
custom_print <- function(res) {
  print(res)
  
  # Extract effect size and convert
  f2 <- res$f2
  R2 <- f2 / (1 + f2)
  
  # Clean output
  cat("\n")
  cat(sprintf("Equivalent R2: %.6f\n", R2))
  cat("\n")
}

```

# AIM 1

-   To identify whether DNA methylation (DNAme) at birth is associated with the life’s essential 8 (LE8) cardiovascular health score components in childhood. (n = 158)

-   From the KEGG CVD, endocrine, and immune pathways we have 1729 genes of interest.

-   1479 of these genes can be found in the autosomal 450k probes n = 32550.

## Hypothesis 1.1

We will examine whether DNA methylation at candidate CpG sites is associated with childhood Life’s Essential 8 (LE8) biological factors score, which includes body mass index, blood lipids, glucose, and blood pressure. The outcome will be the continuous LE8 biological factors score, and the predictor will be DNA methylation level at each CpG site, represented as M-values. For each CpG site (n = 32,550), we will fit a separate multivariable linear regression model adjusting for prespecified covariates. False discovery rate (FDR) adjustment will be used to account for multiple testing across CpG sites.

CpG sites associated with LE8 biological factors score will be annotated to nearby genes, and pathway enrichment analysis will be conducted to determine whether associated genes are enriched in biological pathways related to inflammation and cardiometabolic disease development.

Power calculations were conducted for the CpG-specific multivariable linear regression models using a conservative Bonferroni-adjusted significance level of α = 0.05/32,550 to account for testing across candidate CpG sites. Each model evaluated the incremental contribution of DNA methylation M-value at a given CpG beyond 8 prespecified covariates. At 80% power, this analysis can detect an effect corresponding to Cohen’s f² = 0.219, equivalent to a partial R² of 0.179. This corresponds to a medium-to-large effect size according to Cohen’s convention. Because the planned primary EWAS inference will use FDR adjustment, this Bonferroni-based calculation should be interpreted as a conservative estimate of detectable effect size.

From WebPower (model DNAme, n = 158, alpha = 0.05/n_cpgs, power = 80%):
```{r}
n_cpgs <- 32550
alpha_bonf <- 0.05 / n_cpgs
n_covars <- 8

res <- wp.regression(
  n = 158,      
  p1 = n_covars + 1,  # DNAme + covariates
  p2 = n_covars,
  f2 = NULL,
  alpha = alpha_bonf,
  power = 0.80
)

custom_print(res)
```
## Hypothesis 1.2 ???? 

DNAme at birth partially mediates the association between maternal prepregnancy BMI and poor cardiovascular health in childhood

```{r}

```

# AIM 2

To determine whether infant adiposity is a risk factor of poor cardiovascular health in childhood.

## Hypothesis 2.1 (BOTH 5 months?)

We will examine whether percent fat mass at 5 months is associated with childhood Life’s Essential 8 (LE8) cardiovascular health score among participants with available follow-up data (n = 263). The primary outcome will be the continuous LE8 score, and the primary predictor will be percent fat mass at 5 months. We will use multivariable linear regression adjusting for 8 prespecified covariates. Secondary analyses will examine individual LE8 components as exploratory outcomes.

Power calculations were conducted for the primary multivariable linear regression analysis using n = 263 participants and α = 0.05. The analysis evaluated the incremental contribution of percent fat mass beyond 8 prespecified covariates. We will achieve 80% power to detect an effect corresponding to Cohen’s f² = 0.030, equivalent to a partial R² of 0.029, representing a small effect size according to Cohen’s conventions.

From WebPower (model percent fat mass, n = 263, alpha = 0.05, power = 80%):

```{r}
alpha = 0.05
n_covars = 8

res <- wp.regression(
  n = 263,      
  p1 = n_covars + 1,  # percent fat mass + covariates
  p2 = n_covars,
  f2 = NULL,
  alpha = alpha,
  power = 0.80
)

custom_print(res)
```

## Hypothesis 2.2

We will examine whether rapid weight gain between 0 and 5 months is associated with childhood Life’s Essential 8 (LE8) cardiovascular health score among participants with available follow-up data (n = 273). The primary outcome will be the continuous LE8 score, and the primary predictor will be rapid weight gain. We will use multivariable linear regression adjusting for 8 prespecified covariates. Secondary analyses will examine individual LE8 components as exploratory outcomes.

Power calculations were conducted for the primary multivariable linear regression analysis using n = 263 participants and α = 0.05. The analysis evaluated the incremental contribution of percent fat mass beyond 8 prespecified covariates. We will achieve 80% power to detect an effect corresponding to Cohen’s f² = 0.029, equivalent to a partial R² of 0.028, representing a small effect size according to Cohen’s conventions.

From WebPower (model rapid weight gain, n = 273, alpha = 0.05, power = 80%):

```{r}
alpha = 0.05
n_covars = 8

res <- wp.regression(
  n = 273,      
  p1 = n_covars + 1,  # rapid weight gain + covariates
  p2 = n_covars,
  f2 = NULL,
  alpha = alpha,
  power = 0.80
)

custom_print(res)
```

# AIM 3

To determine the association between LE8 cardiovascular health score and markers of endothelial dysfunction and inflammation among children aged 5 years.

## Hypothesis 3.1

We will examine whether childhood Life’s Essential 8 (LE8) cardiovascular health score is associated with circulating markers of endothelial dysfunction among participants with available data (n = 347). The outcomes will be circulating levels of ICAM, VCAM, E-selectin, and P-selectin, which will be log-transformed if needed. The primary predictor will be continuous LE8 score. Separate multivariable linear regression models will be fit for each endothelial dysfunction marker, adjusting for 8 prespecified covariates.

Power calculations were conducted for multivariable linear regression using n = 347 participants and 8 prespecified covariates. Because four endothelial dysfunction markers will be examined, we used a Bonferroni-adjusted significance level of α = 0.05/4. This likely provides a conservative estimate relative to the planned FDR-adjusted analyses.

We will achieve 80% power to detect an incremental effect of LE8 corresponding to Cohen’s f² = 0.032, equivalent to a partial R² of 0.031, representing a small effect size according to Cohen’s conventions.

From WebPower (model markers of endothelial dysfunction, n = 347, alpha = 0.05/4, power = 80%):

```{r}
alpha_bonf <- 0.05 / 4
n_covars = 8

res <- wp.regression(
  n = 347,      
  p1 = n_covars + 1,  # marker + covariates
  p2 = n_covars,
  f2 = NULL,
  alpha = alpha_bonf,
  power = 0.80
)

custom_print(res)
```

## Hypothesis 3.2

We will examine whether childhood Life’s Essential 8 (LE8) cardiovascular health score is associated with levels of inflammatory markers among participants with available data (n = 222). The outcomes will be CRP and IL6, which will be log-transformed if needed. The primary predictor will be continuous LE8 score. Separate multivariable linear regression models will be fit for each inflammatory marker, adjusting for 8 prespecified covariates.

Power calculations were conducted for multivariable linear regression using n = 222 participants and 8 prespecified covariates. Because two inflammatory markers will be examined, we used a Bonferroni-adjusted significance level of α = 0.05/2. This likely provides a conservative estimate relative to the planned FDR-adjusted analyses.

We will achieve 80% power to detect an incremental effect of LE8 corresponding to Cohen’s f² = 0.043, equivalent to a partial R² of 0.042, representing a small effect size according to Cohen’s conventions.

From WebPower (model markers of endothelial dysfunction, n = 222, alpha = 0.05/2, power = 80%):

```{r}
alpha_bonf <- 0.05 / 2
n_covars = 8

res <- wp.regression(
  n = 222,      
  p1 = n_covars + 1,  # marker + covariates
  p2 = n_covars,
  f2 = NULL,
  alpha = alpha_bonf,
  power = 0.80
)

custom_print(res)
```

```{r, include = FALSE}
sessionInfo()
```
