---
title: "LE8 Calculations"
format:
  html:
    toc: true
    toc-depth: 5
    html-math-method: katex
    embed-resources: true
    smooth-scroll: true
editor: visual
execute:
  echo: true
  include: true
---

# Set up

```{r set-up}
library(rio)
library(tidyverse)
library(ggpubr)
library(gtsummary)
library(gt)
library(GGally)
library(performance)
library(broom)
library(reactable)

source("~/Documents/GitHub/r_helper_scripts/correlation_functions.R")
source("~/Documents/GitHub/r_helper_scripts/two_group_test_functions.R")
```

# Function

```{r}
distribution_check <- function(df,
                               var,
                               numerical_summ = TRUE,
                               hist = TRUE,
                               density = TRUE) {
  out <- list()
  
  if (numerical_summ) {
    out$numeric_summary <- df %>%
      dplyr::select({{ var }}) %>%
      dlookr::diagnose_numeric()
    
    # explicitly print the summary so it shows in the chunk
    print(out$numeric_summary)
  }
  
  if (hist) {
    out$histogram <- df %>%
      dplyr::select({{ var }}) %>%
      DataExplorer::plot_histogram(
        ggtheme = ggplot2::theme_minimal(),
        geom_histogram_args = list(fill = "steelblue", color = "white")
      )
  }
  
  if (density) {
    out$density <- df %>%
      dplyr::select({{ var }}) %>%
      DataExplorer::plot_density(
        ggtheme = ggplot2::theme_minimal(),
        geom_density_args = list(
          color = "steelblue",
          fill = "steelblue",
          alpha = 0.4
        )
      )
  }
  
  # still return it (for optional later use) but without printing the list
  invisible(out)
}

run_spearman_grid <- function(data, clinical_vars, protein_vars) {
  
  cor_results <- expand_grid(
    clinical_var = clinical_vars,
    protein_var = protein_vars
  ) %>%
    mutate(
      cor_test = map2(
        clinical_var,
        protein_var,
        ~ cor.test(
          data[[.x]],
          data[[.y]],
          method = "spearman",
          use = "complete.obs",
          exact = FALSE
        )
      ),
      cor_tidy = map(cor_test, tidy)
    ) %>%
    select(-cor_test) %>%
    unnest(cor_tidy) %>%
    rename(
      spearman_rho = estimate,
      p_value = p.value
    ) %>%
    mutate(
      p_adj_fdr = p.adjust(p_value, method = "fdr")
    ) %>%
    arrange(p_value)
  
  return(cor_results)
}
```

# Wide Data

## PIDs

```{r}
pids_le8_cb <- read.csv("data/PIDS/p0010_ShakyaEvaniya_LE8_CB.csv")
dim(pids_le8_cb)

pids_le8_pfm <- read.csv("data/PIDS/p0010_ShakyaEvaniya_LE8_PFM.csv")

pids_le8_crp_il6 <-read.csv("data/PIDS/p0010_ShakyaEvaniya_LE8_CRP_IL6.csv")

all_pids <- unique(c(pids_le8_cb$PID, pids_le8_pfm$PID, pids_le8_crp_il6$PID))
length(all_pids)
```

```{r}
wide_df <- import("data/LE8 biological_health factors score /df_merged_HS_forAHAgrant.csv")

subset_df <- wide_df %>%
  dplyr::mutate(
    age_months = ageyrs_HSIIVisit * 12,
    ResultNonHDLCholesterol = ResultCholesterol - ResultHDL,
    height_m_HSIIVisit  = sqrt(wtkg_HSIIVisit / bmi_HSIIVisit),
    height_cm_HSIIVisit = height_m_HSIIVisit * 100,
    sex = factor(
      case_when(
        infant_sex == 1 ~ "female",
        infant_sex == 2 ~ "male",
        TRUE ~ NA_character_
      ),
      levels = c("female", "male")
    ),
  ) %>%
  dplyr::select(
    pid,
    childID,
    sex,
    bmi_HSIIVisit,
    age_months,
    ageyrs_HSIIVisit,
    wtkg_HSIIVisit,
    height_cm_HSIIVisit,
    # methylation_ind_HSIIVisit,
    ResultGlucose,
    ResultCholesterol,
    ResultHDL,
    ResultNonHDLCholesterol,
    X_hs2_pdf_bpsys,
    X_hs2_pdf_bpdia,
    ipv4_pp_fm_pct,
    `IL_6_hs___pg_mL`,
    `CRP_hs__mg_L`
  )  %>%
  filter(pid %in% all_pids)

score_df <- subset_df
```

## BMI Score

TODO: NA values for very underweight

-   5th to \<85th percentile = `100`

-   85th to \<95th percentile = `70`

-   ≥95th percentile = `30`

-   120% of the 95th percentile to \<140% of the 95th percentile = `15`

-   ≥140% of the 95th percentile = `0`

```{r}
library(cdcanthro)

summary(score_df$bmi_HSIIVisit)

bmi_df <- score_df %>%
  filter(
    !is.na(bmi_HSIIVisit)
  )

bmi_results <- cdcanthro(
  data = bmi_df,
  age = age_months,
  wt = wtkg_HSIIVisit,
  ht = height_cm_HSIIVisit,
  bmi = bmi_HSIIVisit
) 

bmi_score <- bmi_df %>%
  dplyr::left_join(bmi_results %>% select(pid, bmip, p50, p95, bmip95), by = "pid") %>%
  dplyr::mutate(
    BMI_Score = case_when(
      bmip >= 5  & bmip < 85 ~ 100,
      bmip >= 85 & bmip < 95 ~ 70,
      bmip >= 95 & bmip95 < 120 ~ 30,
      bmip95 >= 120 & bmip95 < 140 ~ 15,
      bmip95 >= 140 ~ 0,
      TRUE ~ NA_real_
    )
  ) 

# see <- bmi_score %>%
#   select(pid, bmi_HSIIVisit, bmip, bmip95, p50, p95, BMI_Score)

```

## Fasting glucose score

TODO: There is one individual with \> 125 mg/dL

-   Non-diabetic with values \<100 mg/dL = `100`

-   Non-diabetic with values 100 to \<125 mg/dL = `60`

```{r}
summary(score_df$ResultGlucose)

gluc_score <- score_df %>%
  mutate(
    Fasting_Glucose_Score = case_when(
      ResultGlucose < 100 ~ 100,
      ResultGlucose >= 100 & ResultGlucose < 125 ~ 60,
      ResultGlucose >= 125 ~ 0,
      TRUE ~ NA_real_
    ),
    Fasting_Glucose_Flag = case_when(
      ResultGlucose >= 125 ~ "ERROR: glucose >=125",
      TRUE ~ NA_character_
    )
  )


```

## Non-HDL cholesterol score

-   \<100 mg/dL = `100`

-   100 to \<120 mg/dL = `60`

-   120 to \<145 mg/dL = `40`

-   145 to \<190 mg/dL = `20`

-   ≥190 mg/dL = `0`

```{r}
summary(score_df$ResultNonHDLCholesterol)

nonhdl_score <- score_df %>%
  mutate(
    NonHDL_Cholesterol_Score = case_when(
      ResultNonHDLCholesterol < 100 ~ 100,
      ResultNonHDLCholesterol >= 100 & ResultNonHDLCholesterol < 120 ~ 60,
      ResultNonHDLCholesterol >= 120 & ResultNonHDLCholesterol < 145 ~ 40,
      ResultNonHDLCholesterol >= 145 & ResultNonHDLCholesterol < 190 ~ 20,
      ResultNonHDLCholesterol >= 190 ~ 0,
      TRUE ~ NA_real_
    )
  )
```

## SBP (bpdia) and DBP (bpdia) score

-   \<90th percentile of age, sex, and height = `100`

-   SBP and DBP 90th to \<95th percentile = `75`

-   95th percentile to 95th percentile + 12 mmHg = `50`

-   95th percentile + 12 mmHg to \<95th percentile + 30 mmHg = `25`

-   ≥95th percentile + 30 mmHg = `0`

```{r}
library(pedbp)

pedbp_df <- score_df %>%
  mutate(
    male = case_when(
      sex == "male" ~ 1,
      sex == "female" ~ 0,
      TRUE ~ NA_real_
    )
  ) %>%
  dplyr::select(pid,
                childID,
                age_months,
                male,
                height_cm = height_cm_HSIIVisit,
                X_hs2_pdf_bpsys,
                X_hs2_pdf_bpdia) %>%
  filter(
    !is.na(X_hs2_pdf_bpsys),
    !is.na(X_hs2_pdf_bpdia),
    !is.na(male),
    !is.na(age_months),
    !is.na(height_cm)
  )

summary(pedbp_df$X_hs2_pdf_bpsys)  # SBP / systolic
summary(pedbp_df$X_hs2_pdf_bpdia)  # DBP / diastolic
summary(pedbp_df$age_months)  # DBP / diastolic
summary(pedbp_df$height_cm)  # DBP / diastolic


# bp_pct_results <- p_bp(
#   q_sbp = pedbp_df$X_hs2_pdf_bpsys,
#   q_dbp = pedbp_df$X_hs2_pdf_bpdia,
#   male   = pedbp_df$male,
#   age    = pedbp_df$age_months,
#   height = pedbp_df$height_cm,
#   source = "martin2022"
# )

# score_df <- pedbp_df %>%
#   mutate(
#     sbp_p = bp_pct_results$sbp_p,
#     dbp_p = bp_pct_results$dbp_p,
#     SBP_percentile = sbp_p * 100,
#     DBP_percentile = dbp_p * 100
#   )


bp_90 <- q_bp(
  p_sbp = 0.90,
  p_dbp = 0.90,
  male   = pedbp_df$male,
  age    = pedbp_df$age_months,
  height = pedbp_df$height_cm,
  source = "martin2022"
)

bp_95 <- q_bp(
  p_sbp = 0.95,
  p_dbp = 0.95,
  male   = pedbp_df$male,
  age    = pedbp_df$age_months,
  height = pedbp_df$height_cm,
  source = "martin2022"
)

pedbp_df <- pedbp_df %>%
  mutate(
    SBP_90 = bp_90$sbp,
    DBP_90 = bp_90$dbp,
    SBP_95 = bp_95$sbp,
    DBP_95 = bp_95$dbp
  )

bp_score <- pedbp_df %>%
  mutate(
    Blood_Pressure_Score = case_when(
      is.na(X_hs2_pdf_bpsys) | is.na(X_hs2_pdf_bpdia) |
        is.na(SBP_90) | is.na(DBP_90) |
        is.na(SBP_95) | is.na(DBP_95) ~ NA_real_,
      
      X_hs2_pdf_bpsys >= SBP_95 + 30 |
        X_hs2_pdf_bpdia >= DBP_95 + 30 ~ 0,
      
      X_hs2_pdf_bpsys >= SBP_95 + 12 |
        X_hs2_pdf_bpdia >= DBP_95 + 12 ~ 25,
      
      X_hs2_pdf_bpsys >= SBP_95 |
        X_hs2_pdf_bpdia >= DBP_95 ~ 50,
      
      X_hs2_pdf_bpsys >= SBP_90 |
        X_hs2_pdf_bpdia >= DBP_90 ~ 75,
      
      X_hs2_pdf_bpsys < SBP_90 &
        X_hs2_pdf_bpdia < DBP_90 ~ 100,
      
      TRUE ~ NA_real_
    )
  )


```

# Table for LE8 Biological

1.1, 2.1, 3.2 Mean and SD p-value (see if different) - maybe later

## Join data

```{r}
bmi_score1 <- bmi_score %>% select(pid, BMI_Score)

gluc_score1  <- gluc_score %>% select(pid, Fasting_Glucose_Score)

nonhdl_score1 <- nonhdl_score  %>% select(pid, NonHDL_Cholesterol_Score)

bp_score1 <- bp_score  %>% select(pid, Blood_Pressure_Score)

score_df_final <- score_df %>%
  left_join(bmi_score1, by = "pid") %>%
  left_join(gluc_score1, by = "pid") %>%
  left_join(nonhdl_score1, by = "pid") %>%
  left_join(bp_score1, by = "pid") %>%
  mutate(
    LE_biological_score = (
      BMI_Score + Fasting_Glucose_Score + NonHDL_Cholesterol_Score + Blood_Pressure_Score
    ) / 4,
    LE_biological_score_categorical = case_when(
      LE_biological_score < 50 ~ "Low",
      LE_biological_score >= 50 &
        LE_biological_score < 80 ~ "Moderate to High",
      LE_biological_score >= 80 &
        LE_biological_score < 100 ~ "Moderate to High",
      LE_biological_score == 100 ~ "Optimal",
      TRUE ~ NA_character_
    ),
    LE_biological_score_binary = case_when(
      LE_biological_score_categorical ==  "Moderate to High" ~ 1,
      LE_biological_score_categorical ==  "Optimal" ~ 0,
      TRUE ~ NA_real_
    )
  )


```

## Subsets

```{r}
pids_le8_cb_score <- score_df_final %>%
  filter(pid %in% pids_le8_cb$PID) %>%
  mutate(hypothesis = "Hypothesis 1.1") %>%
  filter(!is.na(LE_biological_score))
dim(pids_le8_cb_score) #147  17

pids_le8_pfm_score <- score_df_final %>%
  filter(pid %in% pids_le8_pfm$PID) %>%
  mutate(hypothesis = "Hypothesis 2.1") %>%
  filter(!is.na(LE_biological_score))
dim(pids_le8_pfm_score) #250  17

pids_le8_crp_il6_score <- score_df_final %>%
  filter(pid %in% pids_le8_crp_il6$PID) %>%
  mutate(hypothesis = "Hypothesis 3.2") %>%
  filter(!is.na(LE_biological_score))
dim(pids_le8_crp_il6_score) #303  17

full_table_le8 <- rbind(pids_le8_cb_score,
                        pids_le8_pfm_score,
                        pids_le8_crp_il6_score)
```

## Table 1

```{r}
score_vars <- c(
  "LE_biological_score",
  "BMI_Score",
  "Fasting_Glucose_Score",
  "NonHDL_Cholesterol_Score",
  "Blood_Pressure_Score"
)

table1 <- full_table_le8 %>%
  select(hypothesis, all_of(score_vars)) %>%
  mutate(across(all_of(score_vars), as.numeric)) %>%
  tbl_summary(
    by = hypothesis,
    type = all_of(score_vars) ~ "continuous",
    statistic = all_of(score_vars) ~ "{mean} ({sd})",
    digits = all_of(score_vars) ~ 1,
    missing = "ifany",
    label = list(
      LE_biological_score ~ "LE8 Biological",
      BMI_Score ~ "BMI",
      Fasting_Glucose_Score ~ "Fasting Glucose",
      NonHDL_Cholesterol_Score ~ "Non-HDL Cholesterol",
      Blood_Pressure_Score ~ "Blood Pressure"
    )
  ) %>%
  add_p(test = all_of(score_vars) ~ "aov") %>%
  modify_header(label ~ "**Score**", p.value ~ "**p-value**") %>%
  modify_fmt_fun(p.value ~ function(x)
    style_pvalue(x, digits = 3)) %>%
  bold_labels() %>%
  as_gt() %>%
  gt::tab_header(title = "LE8 Biological Scores by Hypothesis")
  
gtsave(table1, "table1_LE8.pdf")

table1


```

## Distribution

```{r}
distribution_check(score_df_final, BMI_Score)
# distribution_check(score_df_final, Fasting_Glucose_Score)
distribution_check(score_df_final, NonHDL_Cholesterol_Score)
distribution_check(score_df_final, Blood_Pressure_Score)
distribution_check(score_df_final, LE_biological_score)
```

# Hypotheses

## 1.1 LE8 vs. Cpgs

n = 158 - lowest priority

(n = 158) Spearman - cpgs ... 12k ones ... adjustment fdr - top 50??? try LE8 or the 4 components separately

```{r}
healthy_start_Ms_12k <- readRDS("healthy_start_Ms_filtered_to_probes_for_genes_of_interest.rds")
dim(healthy_start_Ms_12k) # 588 11300

cpgs_list <- names(healthy_start_Ms_12k)[-1]

cpgs_le8_score <- score_df_final %>%
  filter(childID %in% healthy_start_Ms_12k$pid) %>%
  select(
    pid,
    childID,
    BMI_Score,
    Fasting_Glucose_Score,
    NonHDL_Cholesterol_Score,
    Blood_Pressure_Score,
    LE_biological_score
  ) %>%
  left_join(healthy_start_Ms_12k, by = c("childID" = "pid"))

dim(cpgs_le8_score) #158 11306
```

### LE8 vs. Cpgs

```{r}
le8_results <- run_spearman_grid(
  data = cpgs_le8_score,
  clinical_vars = c("LE_biological_score"),
  protein_vars = cpgs_list
)

reactable(head(le8_results))
```

### BMI score vs. Cpgs

```{r}
bmi_score_results <- run_spearman_grid(
  data = cpgs_le8_score,
  clinical_vars = c("BMI_Score"),
  protein_vars = cpgs_list
)

reactable(head(bmi_score_results))
```

### Fasting_Glucose_Score vs. Cpgs

```{r}
gluc_score_results <- run_spearman_grid(
  data = cpgs_le8_score,
  clinical_vars = c("Fasting_Glucose_Score"),
  protein_vars = cpgs_list
)

reactable(head(gluc_score_results))
```

### NonHDL_Cholesterol_Score vs. Cpgs

```{r}
nonhdl_score_results <- run_spearman_grid(
  data = cpgs_le8_score,
  clinical_vars = c("NonHDL_Cholesterol_Score"),
  protein_vars = cpgs_list
)

reactable(head(nonhdl_score_results))
```

### Blood_Pressure_Score vs. Cpgs

```{r}
bp_score_results <- run_spearman_grid(
  data = cpgs_le8_score,
  clinical_vars = c("Blood_Pressure_Score"),
  protein_vars = cpgs_list
)

reactable(head(bp_score_results))
```

## 1.3 LE8 vs. EAA

n = 147 \### Correlation Matrix

```{r, warning = FALSE, message = FALSE, fig.width=20, fig.height = 20}
clock_df <- readRDS("~/Documents/GitHub/healthy_start/data/clock_data/select_clock_results.rds")

pids_le8_cb_score_clock <- pids_le8_cb_score %>%
  mutate(childID = as.character(childID)) %>%
  left_join(clock_df, by = "childID")
dim(pids_le8_cb_score_clock)

pids_le8_cb_score_clock_matrix <- pids_le8_cb_score_clock %>%
  select(
    sex,
    BMI_Score,
    Fasting_Glucose_Score,
    NonHDL_Cholesterol_Score,
    Blood_Pressure_Score,
    LE_biological_score,
    contains("Knight"),
    contains("Bohlin"),
    contains("Horvath")
  )
  
corr_matrix_plot(pids_le8_cb_score_clock_matrix, "sex", method = "spearman", 2)
```

### LE8 Cont. vs. EAA

Spearman correlations

```{r}
plot_correlation_by_group(
  "",
  pids_le8_cb_score_clock_matrix,
  "ageAcc2.Bohlin",
  "Blood_Pressure_Score",
  "ageAcc2.Bohlin",
  "Blood_Pressure_Score",
  method = "spearman",
  group_var = "sex"
)

plot_correlation_by_group(
  "",
  pids_le8_cb_score_clock_matrix,
  "ageAcc.Bohlin",
  "Blood_Pressure_Score",
  "ageAcc.Bohlin",
  "Blood_Pressure_Score",
  method = "spearman",
  group_var = "sex"
)
```

### LE8 Cat. EAA

t-test of LE_biological_score_categorical vs. EAA

```{r, fig.width = 8, fig.height = 6}
results <- run_unpaired_comparison(
  pids_le8_cb_score_clock,
  "LE_biological_score_categorical",
  "ageAcc.Bohlin",
  group_name = NULL,
  outcome_name = NULL,
  ttest_pval_threshold = 0.05,
  shapiro_pval_threshold = 0.05
) 

results <- run_unpaired_comparison(
  pids_le8_cb_score_clock,
  "LE_biological_score_categorical",
  "ageAcc2.Bohlin",
  group_name = NULL,
  outcome_name = NULL,
  ttest_pval_threshold = 0.05,
  shapiro_pval_threshold = 0.05
) 

results <- run_unpaired_comparison(
  pids_le8_cb_score_clock,
  "LE_biological_score_categorical",
  "ageAcc.Knight",
  group_name = NULL,
  outcome_name = NULL,
  ttest_pval_threshold = 0.05,
  shapiro_pval_threshold = 0.05
) 

results <- run_unpaired_comparison(
  pids_le8_cb_score_clock,
  "LE_biological_score_categorical",
  "ageAcc2.Knight",
  group_name = NULL,
  outcome_name = NULL,
  ttest_pval_threshold = 0.05,
  shapiro_pval_threshold = 0.05
) 
```

## 2.1 Fat Mass vs. Biological LE8 n = 250

We will examine whether percent fat mass at 5 months is associated with childhood Life’s Essential 8 (LE8) cardiovascular health score among participants with available follow-up data (n = 263).

Correlation `ipv4_pp_fm_pct`

### Table 1

```{r}
table1_fm_sex <- pids_le8_pfm_score %>%
  select(sex, ipv4_pp_fm_pct) %>%
  tbl_summary(
    by = sex,
    missing = "ifany",
    label = list(
      ipv4_pp_fm_pct ~ "Fat Mass"
    )
  ) %>%
  add_p() %>%
  bold_labels() %>%
  as_gt() %>%
  gt::tab_header(title = "Fat Mass by Sex")

table1_fm_sex

table1_fm_le8cat <- pids_le8_pfm_score %>%
  select(LE_biological_score_categorical, ipv4_pp_fm_pct) %>%
  tbl_summary(
    by = LE_biological_score_categorical,
    missing = "ifany",
    label = list(
      ipv4_pp_fm_pct ~ "Fat Mass"
    )
  ) %>%
  add_p() %>%
  bold_labels() %>%
  as_gt() %>%
  gt::tab_header(title = "Fat Mass by LE8 Biological Score")

table1_fm_le8cat
```

### LE8 Cont. vs. fm

```{r}
dim(pids_le8_pfm_score) #250  23

plot_correlation(
  "",
  pids_le8_pfm_score,
  "ipv4_pp_fm_pct",
  "LE_biological_score",
  "ipv4_pp_fm_pct",
  "LE_biological_score",
  method = "spearman",
  show_ci = FALSE
)

plot_correlation_by_group(
  "",
  pids_le8_pfm_score,
  "ipv4_pp_fm_pct",
  "LE_biological_score",
  "ipv4_pp_fm_pct",
  "LE_biological_score",
  method = "spearman",
  group_var = "sex"
)

```

### LE8 Cat. fm

```{r, fig.width = 8, fig.height = 6}
results <- run_unpaired_comparison(
  pids_le8_pfm_score,
  "LE_biological_score_categorical",
  "ipv4_pp_fm_pct",
  group_name = NULL,
  outcome_name = NULL,
  ttest_pval_threshold = 0.05,
  shapiro_pval_threshold = 0.05
) 
```

### LE8 Binary fm

Logistic Regression

```{r}
model <- glm(
  LE_biological_score_binary ~ ipv4_pp_fm_pct + sex,
  data = pids_le8_pfm_score,
  family = "binomial"
)
summary(model)
check_model(model)

model_interaction <- glm(
  LE_biological_score_binary ~ ipv4_pp_fm_pct * sex,
  data = pids_le8_pfm_score,
  family = "binomial"
)
summary(model_interaction)
check_model(model_interaction)
```

## 3.2 CRP/IL6 vs. Biological LE8 n = 303

We will examine whether childhood Life’s Essential 8 (LE8) cardiovascular health score is associated with levels of inflammatory markers among participants with available data (n = 322). The outcomes will be CRP and IL6, which will be log-transformed if needed

IL-6 values reported as below the lower limit of detection (\<0.16) were treated as left-censored and assigned a value of 0.16/√2 prior to log transformation. IL-6 was then analyzed as a continuous log-transformed biomarker. A flag for values below the detection limit was retained for descriptive purposes.

### Clean

```{r}
pids_le8_crp_il6_score <- pids_le8_crp_il6_score %>%
  mutate(
    IL6_clean = IL_6_hs___pg_mL,
    IL6_clean = case_when(
      IL6_clean == "Less than 0.16" ~ 0.16 / sqrt(2),
      IL6_clean == "rpt x10" ~ NA,
      TRUE ~ parse_number(IL6_clean)
    ),
    log_IL6_clean = log(IL6_clean),
    IL6_tertile = ntile(IL6_clean, 3),
    IL6_tertile = factor(
      IL6_tertile,
      levels = 1:3,
      labels = c("T1: Low IL-6", "T2: Medium IL-6", "T3: High IL-6")
    ),
    CRP_clean = IL_6_hs___pg_mL,
    CRP_clean = case_when(
      CRP_clean == "Less than 0.2" ~ 0.2 / sqrt(2),
      CRP_clean == "<0.2"  ~ 0.2 / sqrt(2),
      TRUE ~ parse_number(CRP_clean)
    ),
    log_CRP_clean = log(CRP_clean),
    CRP_category = case_when(
      CRP_clean < 0.3 ~ "Normal to Minor elevation",
      CRP_clean >= 0.3 & CRP_clean < 1.0 ~ "Normal to Minor elevation",
      CRP_clean >= 1.0 & CRP_clean <= 10.0 ~ "Moderate to Marked elevation",
      CRP_clean > 10.0 ~ "Moderate to Marked elevation",
      TRUE ~ NA_character_
    ),
    CRP_category = factor(
      CRP_category,
      levels = c(
        "Normal to Minor elevation",
        "Moderate to Marked elevation"
        # "Normal",
        # "Minor elevation",
        # "Moderate elevation",
        # "Marked elevation"
      ),
      ordered = TRUE
    )
  )

dim(pids_le8_crp_il6_score)
```

### Table 1

```{r}
table1_crp_il6 <- pids_le8_crp_il6_score %>%
  select(LE_biological_score_categorical, CRP_category, IL6_tertile) %>%
  tbl_summary(
    by = LE_biological_score_categorical,
    missing = "ifany",
    label = list(
      CRP_category ~ "CRP Categorical",
      IL6_tertile ~ "IL6 Tertile"
    )
  ) %>%
  add_p() %>%
  bold_labels() %>%
  as_gt() %>%
  gt::tab_header(title = "CRP and IL6 by LE8 Biological Scores Category")

table1_crp_il6

## Other Table
score_vars <- c(
  "LE_biological_score",
  "BMI_Score",
  "Fasting_Glucose_Score",
  "NonHDL_Cholesterol_Score",
  "Blood_Pressure_Score"
)

table_crp <- pids_le8_crp_il6_score %>%
  select(CRP_category, all_of(score_vars)) %>%
  mutate(across(all_of(score_vars), as.numeric)) %>%
  tbl_summary(
    by = CRP_category,
    type = all_of(score_vars) ~ "continuous",
    statistic = all_of(score_vars) ~ "{mean} ({sd})",
    digits = all_of(score_vars) ~ 1,
    missing = "ifany",
    label = list(
      LE_biological_score ~ "LE8 Biological",
      BMI_Score ~ "BMI",
      Fasting_Glucose_Score ~ "Fasting Glucose",
      NonHDL_Cholesterol_Score ~ "Non-HDL Cholesterol",
      Blood_Pressure_Score ~ "Blood Pressure"
    )
  ) %>%
  add_p(test = all_of(score_vars) ~ "aov") %>%
  modify_header(label ~ "**Score**", p.value ~ "**p-value**") %>%
  modify_fmt_fun(p.value ~ function(x)
    style_pvalue(x, digits = 3)) %>%
  bold_labels() %>%
  as_gt() %>%
  gt::tab_header(title = "LE8 Biological Scores by CRP Category")

table_crp

table_il6 <- pids_le8_crp_il6_score %>%
  select(IL6_tertile, all_of(score_vars)) %>%
  mutate(across(all_of(score_vars), as.numeric)) %>%
  tbl_summary(
    by = IL6_tertile,
    type = all_of(score_vars) ~ "continuous",
    statistic = all_of(score_vars) ~ "{mean} ({sd})",
    digits = all_of(score_vars) ~ 1,
    missing = "ifany",
    label = list(
      LE_biological_score ~ "LE8 Biological",
      BMI_Score ~ "BMI",
      Fasting_Glucose_Score ~ "Fasting Glucose",
      NonHDL_Cholesterol_Score ~ "Non-HDL Cholesterol",
      Blood_Pressure_Score ~ "Blood Pressure"
    )
  ) %>%
  add_p(test = all_of(score_vars) ~ "aov") %>%
  modify_header(label ~ "**Score**", p.value ~ "**p-value**") %>%
  modify_fmt_fun(p.value ~ function(x)
    style_pvalue(x, digits = 3)) %>%
  bold_labels() %>%
  as_gt() %>%
  gt::tab_header(title = "LE8 Biological Scores by IL-6 Tertile")


table_il6
```

### LE8 Cont. vs. CRP/IL6 Cont.

```{r}
plot_correlation(
  "",
  pids_le8_crp_il6_score,
  "log_CRP_clean",
  "LE_biological_score",
  "log_CRP_clean",
  "LE_biological_score",
  method = "spearman",
  show_ci = FALSE
)

plot_correlation_by_group(
  "",
  pids_le8_crp_il6_score,
  "log_CRP_clean",
  "LE_biological_score",
  "log_CRP_clean",
  "LE_biological_score",
  method = "spearman",
  group_var = "sex"
)

plot_correlation(
  "",
  pids_le8_crp_il6_score,
  "log_IL6_clean",
  "LE_biological_score",
  "log_IL6_clean",
  "LE_biological_score",
  method = "spearman",
  show_ci = FALSE
)

plot_correlation_by_group(
  "",
  pids_le8_crp_il6_score,
  "log_IL6_clean",
  "LE_biological_score",
  "log_IL6_clean",
  "LE_biological_score",
  method = "spearman",
  group_var = "sex"
)
```
