---
title: "Methylation Power Calculations (R)"
date: today
format:
  docx:
    toc: true
    number-sections: true
    syntax-highlighting: github
  pdf: 
    documentclass: article
  html:
    toc: true
    toc-depth: 5
    html-math-method: katex
    embed-resources: true
    smooth-scroll: true
editor: visual
execute:
  echo: true
  include: true
---

# Set up

```{r, include = FALSE}
library(pwr)
library(tidyverse)
library(gt)
```

Aim 1 is (Healthy Start Data): Aim 1: To identify whether DNA methylation (DNAme) at birth is associated with the life’s essential 8 (LE8) cardiovascular health score components in childhood. (n = 158) Hypothesis 1.1: DNAme sites associated with lower LE8 biological factors score (body mass index, blood lipids, glucose, and blood pressure) will be enriched in pathways related to inflammation and development of cardiometabolic disease. Hypothesis 1.2: DNAme at birth partially mediates the association between maternal prepregnancy BMI and poor cardiovascular health in childhood

# Calculate Power

439,532 429,246

Power to look at specific genes for pathway We are powered to EWAS to look at large changes of ...

```{r}
n <- 158
n_covariates <- 12
u <- 1               # testing one CpG term
v <- n - n_covariates - u - 1

# 429246

alpha_nominal <- 0.05
alpha_bonf <- 0.05 / 20000

power_for_partial_r2 <- function(partial_r2, alpha) {
  f2 <- partial_r2 / (1 - partial_r2)
  
  pwr.f2.test(
    u = u,
    v = v,
    f2 = f2,
    sig.level = alpha
  )$power
}

# Define effect size categories
effect_sizes <- tibble(
  category = c("Small", "Moderate low", "Moderate high", "Large low", "Large high"),
  partial_R2 = c(0.05, 0.10, 0.15, 0.20, 0.25)
)

# Calculate power
power_results <- effect_sizes %>%
  mutate(
    nominal_power = power_for_partial_r2(partial_R2, alpha_nominal),
    bonf_power = power_for_partial_r2(partial_R2, alpha_bonf)
  )

power_results
```

## Format

```{r}
format_power <- function(x) {
  case_when(
    x < 0.01 ~ "<1%",
    x > 0.995 ~ ">99%",
    TRUE ~ paste0(round(x * 100), "%")
  )
}

table_power <- tibble(
  `Dataset / Tissue Type` = c(
    "Healthy Start birth DNAm, nominal α = 0.05",
    "Healthy Start birth DNAm, conservative α = 0.05/429,246"
  ),
  `Sample Size` = "N = 158",
  `Small Effect: 5% partial R²` = c(
    format_power(power_results$nominal_power[power_results$partial_R2 == 0.05]),
    format_power(power_results$bonf_power[power_results$partial_R2 == 0.05])
  ),
  `Moderate Effect: 10–15% partial R²` = c(
    paste0(
      format_power(power_results$nominal_power[power_results$partial_R2 == 0.10]),
      "–",
      format_power(power_results$nominal_power[power_results$partial_R2 == 0.15])
    ),
    paste0(
      format_power(power_results$bonf_power[power_results$partial_R2 == 0.10]),
      "–",
      format_power(power_results$bonf_power[power_results$partial_R2 == 0.15])
    )
  ),
  `Large Effect: 20–25% partial R²` = c(
    paste0(
      format_power(power_results$nominal_power[power_results$partial_R2 == 0.20]),
      "–",
      format_power(power_results$nominal_power[power_results$partial_R2 == 0.25])
    ),
    paste0(
      format_power(power_results$bonf_power[power_results$partial_R2 == 0.20]),
      "–",
      format_power(power_results$bonf_power[power_results$partial_R2 == 0.25])
    )
  )
)

table_power
```

## Format GT

```{r}
power_gt <- table_power %>%
  gt() %>%
  tab_header(
    title = md("**Expected Power at Varying Effect Sizes for CpG-level Associations**")
  ) %>%
  tab_source_note(
    source_note = md(
      "Power estimated for multivariable linear regression models testing one CpG term, assuming N = 158 and 10 covariates. Effect sizes are expressed as partial R², representing the additional variance explained by each CpG after covariate adjustment. Expected power at varying effect sizes for association between DNAm at birth and childhood LE8 biological factors score."
    )
  ) %>%
  cols_align(
    align = "center",
    columns = -`Dataset / Tissue Type`
  ) %>%
  cols_align(
    align = "left",
    columns = `Dataset / Tissue Type`
  ) %>%
  tab_options(
    table.font.size = 12,
    heading.title.font.size = 13,
    data_row.padding = px(4)
  )

power_gt
```

# WebPower

```{r}
library(WebPower)
library(dplyr)

n <- 158
n_covariates <- 10

# total predictors in the full model:
# LE8 predictor + covariates
p <- n_covariates + 1

alpha_nominal <- 0.05
alpha_bonf <- 0.05 / 930000  # EPIC v2 conservative threshold

effect_sizes <- tibble(
  effect_size = c("Very small", "Small", "Moderate", "Large"),
  partial_R2 = c(0.005, 0.01, 0.02, 0.05)
) %>%
  mutate(
    f2 = partial_R2 / (1 - partial_R2)
  )

effect_sizes
```

## Hmmm

```{r}
effect_sizes <- effect_sizes %>%
  rowwise() %>%
  mutate(
    power_nominal = wp.regression(
      n = n,
      p1 = p,
      p2 = n_covariates,
      f2 = f2,
      alpha = alpha_nominal
    )$power,
    
    power_bonferroni = wp.regression(
      n = n,
      p1 = p,
      p2 = n_covariates,
      f2 = f2,
      alpha = alpha_bonf
    )$power
  ) %>%
  ungroup()

effect_sizes
```

# Session Info

```{r, include = FALSE}
sessionInfo()
```
