---
title: "LE8 Calculations"
format:
  html:
    toc: true
    toc-depth: 5
    html-math-method: katex
    embed-resources: true
    smooth-scroll: true
editor: visual
execute:
  echo: true
  include: true
---

# Set up

```{r set-up}
library(rio)
library(tidyverse)
library(ggpubr)
source("~/Documents/GitHub/r_helper_scripts/correlation_functions.R")
```

# Wide Data

## PIDs
```{r}
pids_le8_cb <- read.csv("~/Documents/GitHub/hs_power_analysis/PIDS/p0010_ShakyaEvaniya_LE8_CB.csv")
dim(pids_le8_cb)

pids_le8_pfm <- read.csv("~/Documents/GitHub/hs_power_analysis/PIDS/p0010_ShakyaEvaniya_LE8_PFM.csv")

pids_le8_crp_il6 <-read.csv("~/Documents/GitHub/hs_power_analysis/PIDS/p0010_ShakyaEvaniya_LE8_CRP_IL6.csv")

all_pids <- unique(c(pids_le8_cb$PID, pids_le8_pfm$PID, pids_le8_crp_il6$PID))
length(all_pids)
```

```{r}
wide_df <- import("~/Documents/GitHub/hs_power_analysis/LE8 biological_health factors score /df_merged_HS_forAHAgrant.csv")

subset_df <- wide_df %>%
  dplyr::mutate(
    age_months = ageyrs_HSIIVisit * 12,
    ResultNonHDLCholesterol = ResultCholesterol - ResultHDL,
    height_m_HSIIVisit  = sqrt(wtkg_HSIIVisit / bmi_HSIIVisit),
    height_cm_HSIIVisit = height_m_HSIIVisit * 100,
    sex = factor(
      case_when(
        infant_sex == 1 ~ "female",
        infant_sex == 2 ~ "male",
        TRUE ~ NA_character_
      ),
      levels = c("female", "male")
    ),
  ) %>%
  dplyr::select(
    pid,
    childID,
    sex,
    bmi_HSIIVisit,
    age_months,
    ageyrs_HSIIVisit,
    wtkg_HSIIVisit,
    height_cm_HSIIVisit,
    # methylation_ind_HSIIVisit,
    ResultGlucose,
    ResultCholesterol,
    ResultHDL,
    ResultNonHDLCholesterol,
    X_hs2_pdf_bpsys,
    X_hs2_pdf_bpdia,
    ipv4_pp_fm_pct,
    `IL_6_hs___pg_mL`,
    `CRP_hs__mg_L`
  )  %>%
  filter(pid %in% all_pids)

score_df <- subset_df
```

## BMI Score

TODO: NA values for very underweight

-   5th to \<85th percentile = `100`

-   85th to \<95th percentile = `70`

-   ≥95th percentile = `30`

-   120% of the 95th percentile to \<140% of the 95th percentile = `15`

-   ≥140% of the 95th percentile = `0`

```{r}
library(cdcanthro)

summary(score_df$bmi_HSIIVisit)

bmi_df <- score_df %>%
  filter(
    !is.na(bmi_HSIIVisit)
  )

bmi_results <- cdcanthro(
  data = bmi_df,
  age = age_months,
  wt = wtkg_HSIIVisit,
  ht = height_cm_HSIIVisit,
  bmi = bmi_HSIIVisit
) 

bmi_score <- bmi_df %>%
  dplyr::left_join(bmi_results %>% select(pid, bmip, p50, p95, bmip95), by = "pid") %>%
  dplyr::mutate(
    BMI_Score = case_when(
      bmip >= 5  & bmip < 85 ~ 100,
      bmip >= 85 & bmip < 95 ~ 70,
      bmip >= 95 & bmip95 < 120 ~ 30,
      bmip95 >= 120 & bmip95 < 140 ~ 15,
      bmip95 >= 140 ~ 0,
      TRUE ~ NA_real_
    )
  ) 

# see <- bmi_score %>%
#   select(pid, bmi_HSIIVisit, bmip, bmip95, p50, p95, BMI_Score)

```

## Fasting glucose score

TODO: There is one individual with \> 125 mg/dL

-   Non-diabetic with values \<100 mg/dL = `100`

-   Non-diabetic with values 100 to \<125 mg/dL = `60`

```{r}
summary(score_df$ResultGlucose)

gluc_score <- score_df %>%
  mutate(
    Fasting_Glucose_Score = case_when(
      ResultGlucose < 100 ~ 100,
      ResultGlucose >= 100 & ResultGlucose < 125 ~ 60,
      ResultGlucose >= 125 ~ 0,
      TRUE ~ NA_real_
    ),
    Fasting_Glucose_Flag = case_when(
      ResultGlucose >= 125 ~ "ERROR: glucose >=125",
      TRUE ~ NA_character_
    )
  )


```

## Non-HDL cholesterol score

-   \<100 mg/dL = `100`

-   100 to \<120 mg/dL = `60`

-   120 to \<145 mg/dL = `40`

-   145 to \<190 mg/dL = `20`

-   ≥190 mg/dL = `0`

```{r}
summary(score_df$ResultNonHDLCholesterol)

nonhdl_score <- score_df %>%
  mutate(
    NonHDL_Cholesterol_Score = case_when(
      ResultNonHDLCholesterol < 100 ~ 100,
      ResultNonHDLCholesterol >= 100 & ResultNonHDLCholesterol < 120 ~ 60,
      ResultNonHDLCholesterol >= 120 & ResultNonHDLCholesterol < 145 ~ 40,
      ResultNonHDLCholesterol >= 145 & ResultNonHDLCholesterol < 190 ~ 20,
      ResultNonHDLCholesterol >= 190 ~ 0,
      TRUE ~ NA_real_
    )
  )
```

## SBP (bpdia) and DBP (bpdia) score

-   \<90th percentile of age, sex, and height = `100`

-   SBP and DBP 90th to \<95th percentile = `75`

-   95th percentile to 95th percentile + 12 mmHg = `50`

-   95th percentile + 12 mmHg to \<95th percentile + 30 mmHg = `25`

-   ≥95th percentile + 30 mmHg = `0`

```{r}
library(pedbp)

pedbp_df <- score_df %>%
  mutate(
    male = case_when(
      sex == "male" ~ 1,
      sex == "female" ~ 0,
      TRUE ~ NA_real_
    )
  ) %>%
  dplyr::select(pid,
                childID,
                age_months,
                male,
                height_cm = height_cm_HSIIVisit,
                X_hs2_pdf_bpsys,
                X_hs2_pdf_bpdia) %>%
  filter(
    !is.na(X_hs2_pdf_bpsys),
    !is.na(X_hs2_pdf_bpdia),
    !is.na(male),
    !is.na(age_months),
    !is.na(height_cm)
  )

summary(pedbp_df$X_hs2_pdf_bpsys)  # SBP / systolic
summary(pedbp_df$X_hs2_pdf_bpdia)  # DBP / diastolic
summary(pedbp_df$age_months)  # DBP / diastolic
summary(pedbp_df$height_cm)  # DBP / diastolic


# bp_pct_results <- p_bp(
#   q_sbp = pedbp_df$X_hs2_pdf_bpsys,
#   q_dbp = pedbp_df$X_hs2_pdf_bpdia,
#   male   = pedbp_df$male,
#   age    = pedbp_df$age_months,
#   height = pedbp_df$height_cm,
#   source = "martin2022"
# )

# score_df <- pedbp_df %>%
#   mutate(
#     sbp_p = bp_pct_results$sbp_p,
#     dbp_p = bp_pct_results$dbp_p,
#     SBP_percentile = sbp_p * 100,
#     DBP_percentile = dbp_p * 100
#   )


bp_90 <- q_bp(
  p_sbp = 0.90,
  p_dbp = 0.90,
  male   = pedbp_df$male,
  age    = pedbp_df$age_months,
  height = pedbp_df$height_cm,
  source = "martin2022"
)

bp_95 <- q_bp(
  p_sbp = 0.95,
  p_dbp = 0.95,
  male   = pedbp_df$male,
  age    = pedbp_df$age_months,
  height = pedbp_df$height_cm,
  source = "martin2022"
)

score_df <- pedbp_df %>%
  mutate(
    SBP_90 = bp_90$sbp,
    DBP_90 = bp_90$dbp,
    SBP_95 = bp_95$sbp,
    DBP_95 = bp_95$dbp
  )

bp_score <- score_df %>%
  mutate(
    Blood_Pressure_Score = case_when(
      is.na(X_hs2_pdf_bpsys) | is.na(X_hs2_pdf_bpdia) |
        is.na(SBP_90) | is.na(DBP_90) |
        is.na(SBP_95) | is.na(DBP_95) ~ NA_real_,
      
      X_hs2_pdf_bpsys >= SBP_95 + 30 |
        X_hs2_pdf_bpdia >= DBP_95 + 30 ~ 0,
      
      X_hs2_pdf_bpsys >= SBP_95 + 12 |
        X_hs2_pdf_bpdia >= DBP_95 + 12 ~ 25,
      
      X_hs2_pdf_bpsys >= SBP_95 |
        X_hs2_pdf_bpdia >= DBP_95 ~ 50,
      
      X_hs2_pdf_bpsys >= SBP_90 |
        X_hs2_pdf_bpdia >= DBP_90 ~ 75,
      
      X_hs2_pdf_bpsys < SBP_90 &
        X_hs2_pdf_bpdia < DBP_90 ~ 100,
      
      TRUE ~ NA_real_
    )
  )


```

# Table for LE8 Biological
1.1, 2.1, 3.2 
Mean and SD
p-value (see if different) - maybe later 

## Join data
```{r}
bmi_score1 <- bmi_score %>% select(pid, BMI_Score)

gluc_score1  <- gluc_score %>% select(pid, Fasting_Glucose_Score)

nonhdl_score1 <- nonhdl_score  %>% select(pid, NonHDL_Cholesterol_Score)

bp_score1 <- bp_score  %>% select(pid, Blood_Pressure_Score)

score_df_final <- score_df %>%
  left_join(bmi_score1, by = "pid") %>%
  left_join(gluc_score1, by = "pid") %>%
  left_join(nonhdl_score1, by = "pid") %>%
  left_join(bp_score1, by = "pid") 

```

## Subsets 

```{r}
pids_le8_cb_score <- score_df_final %>%
  filter(pid %in% pids_le8_cb$PID) %>%
  mutate(hypothesis = "Hypothesis 1.1") %>%
  mutate(
    LE_biological_score = (
      BMI_Score + Fasting_Glucose_Score + NonHDL_Cholesterol_Score + Blood_Pressure_Score
    ) / 4
  ) %>%
  filter(!is.na(LE_biological_score))
dim(pids_le8_cb_score) #147  17

pids_le8_pfm_score <- score_df_final %>%
  filter(pid %in% pids_le8_pfm$PID) %>%
  mutate(hypothesis = "Hypothesis 2.1") %>%
  mutate(
    LE_biological_score = (
      BMI_Score + Fasting_Glucose_Score + NonHDL_Cholesterol_Score + Blood_Pressure_Score
    ) / 4
  )%>%
  filter(!is.na(LE_biological_score))
dim(pids_le8_pfm_score) #250  17

pids_le8_crp_il6_score <- score_df_final %>%
  filter(pid %in% pids_le8_crp_il6$PID) %>%
  mutate(hypothesis = "Hypothesis 3.2") %>%
  mutate(
    LE_biological_score = (
      BMI_Score + Fasting_Glucose_Score + NonHDL_Cholesterol_Score + Blood_Pressure_Score
    ) / 4
  )%>%
  filter(!is.na(LE_biological_score))
dim(pids_le8_crp_il6_score) #303  17

full_table_le8 <- rbind(pids_le8_cb_score,
                        pids_le8_pfm_score,
                        pids_le8_crp_il6_score)
```

## Table 1
```{r}
library(gtsummary)
library(gt)

score_vars <- c(
  "LE_biological_score",
  "BMI_Score",
  "Fasting_Glucose_Score",
  "NonHDL_Cholesterol_Score",
  "Blood_Pressure_Score"
)

table1 <- full_table_le8 %>%
  select(hypothesis, all_of(score_vars)) %>%
  mutate(across(all_of(score_vars), as.numeric)) %>%
  tbl_summary(
    by = hypothesis,
    type = all_of(score_vars) ~ "continuous",
    statistic = all_of(score_vars) ~ "{mean} ({sd})",
    digits = all_of(score_vars) ~ 1,
    missing = "no",
    label = list(
      LE_biological_score ~ "LE8 Biological",
      BMI_Score ~ "BMI",
      Fasting_Glucose_Score ~ "Fasting Glucose",
      NonHDL_Cholesterol_Score ~ "Non-HDL Cholesterol",
      Blood_Pressure_Score ~ "Blood Pressure"
    )
  ) %>%
  add_p(test = all_of(score_vars) ~ "aov") %>%
  modify_header(label ~ "**Score**", p.value ~ "**p-value**") %>%
  modify_fmt_fun(p.value ~ function(x)
    style_pvalue(x, digits = 3)) %>%
  bold_labels() %>%
  as_gt()
  
gtsave(table1, "table1_LE8.pdf")

table1


```


# Hypotheses 




## 1.1 Biological LE8 n = 158 - lowest priority 
(n = 158)
Spearman - cpgs ... 12k ones ... adjustment fdr - top 50??? try LE8 or the 4 components separately 
```{r}


```

## 1.3 Biological LE8 vs. EAA??? n = 147
Knight & Bohlin, LE8 and components - correlation plots or heat plot? 
```{r}
dim(pids_le8_cb_score)
corr_matrix_plot <- function(df, group_var, method = "spearman", start_index)

```



## 2.1 Fat Mass vs. Biological LE8 n = 250
We will examine whether percent fat mass at 5 months is associated with childhood Life’s Essential 8 (LE8) cardiovascular health score among participants with available follow-up data (n = 263).

Correlation
`ipv4_pp_fm_pct`
```{r}
pids_le8_pfm_score
dim(pids_le8_pfm_score) #250  23

cor.test()

cor1 <- cor.test(
  pids_le8_pfm_score$ipv4_pp_fm_pct,
  pids_le8_pfm_score$LE_biological_score,
  alternative = "two.sided",
  method = "spearman",
  conf.level = 0.95
)

plot_correlation("",pids_le8_pfm_score,"ipv4_pp_fm_pct","LE_biological_score", "ipv4_pp_fm_pct","LE_biological_score", method = "spearman", show_ci = FALSE)


wide_df$ipv4_pp_fm_pct

```


## 3.2 CRP/IL6 vs. Biological LE8 n = 303
We will examine whether childhood Life’s Essential 8 (LE8) cardiovascular health score is associated with levels of inflammatory markers among participants with available data (n = 322). The outcomes will be CRP and IL6, which will be log-transformed if needed

Correlation 
- Log transform CRP and IL6
`IL_6_hs___pg_mL`                 `CRP_hs__mg_L`

```{r}
pids_le8_crp_il6_score <- pids_le8_crp_il6_score %>%
  mutate(
    CRP_hs__mg_L_log = log(CRP_hs__mg_L),
    IL_6_hs___pg_mL_log = log(IL_6_hs___pg_mL)
  )
dim(pids_le8_crp_il6_score)

plot_correlation("",pids_le8_crp_il6_score,"CRP_hs__mg_L","LE_biological_score", "CRP_hs__mg_L","LE_biological_score", method = "spearman", show_ci = FALSE)

plot_correlation("",pids_le8_crp_il6_score,"IL_6_hs___pg_mL","LE_biological_score", "IL_6_hs___pg_mL","LE_biological_score", method = "spearman", show_ci = FALSE)
```


