---
title: "LE8 Calculations"
format:
  html:
    toc: true
    toc-depth: 5
    html-math-method: katex
    embed-resources: true
    smooth-scroll: true
editor: visual
execute:
  echo: true
  include: true
---

# Set up

```{r set-up}
library(rio)
library(tidyverse)
```

# Wide Data

```{r}
wide_df <- import("~/Documents/GitHub/hs_power_analysis/LE8 biological_health factors score /df_merged_HS_forAHAgrant.csv")

subset_df <- wide_df %>%
  dplyr::mutate(
    age_months = ageyrs_HSIIVisit * 12,
    ResultNonHDLCholesterol = ResultCholesterol - ResultHDL,
    height_m_HSIIVisit  = sqrt(wtkg_HSIIVisit / bmi_HSIIVisit),
    height_cm_HSIIVisit = height_m_HSIIVisit * 100,
    sex = factor(
      case_when(
        infant_sex == 1 ~ "female",
        infant_sex == 2 ~ "male",
        TRUE ~ NA_character_
      ),
      levels = c("female", "male")
    ),
  ) %>%
  dplyr::select(
    pid,
    childID,
    sex,
    bmi_HSIIVisit,
    age_months,
    ageyrs_HSIIVisit,
    wtkg_HSIIVisit,
    height_cm_HSIIVisit,
    methylation_ind_HSIIVisit,
    ResultGlucose,
    ResultCholesterol,
    ResultHDL,
    ResultNonHDLCholesterol,
    X_hs2_pdf_bpsys,
    X_hs2_pdf_bpdia
  )  

score_df <- subset_df
```

# BMI Score
TODO: NA values for very underweight

-   5th to \<85th percentile = `100`

-   85th to \<95th percentile = `70`

-   ≥95th percentile = `30`

-   120% of the 95th percentile to \<140% of the 95th percentile = `15`

-   ≥140% of the 95th percentile = `0`

```{r}
library(cdcanthro)

summary(score_df$bmi_HSIIVisit)

bmi_df <- score_df %>%
  filter(
    !is.na(bmi_HSIIVisit)
  )

bmi_results <- cdcanthro(
  data = bmi_df,
  age = age_months,
  wt = wtkg_HSIIVisit,
  ht = height_cm_HSIIVisit,
  bmi = bmi_HSIIVisit
) 

see <- bmi_score %>%
  select(pid, bmi_HSIIVisit, bmip, bmip95, p50, p95, BMI_Score)

bmi_score <- bmi_df %>%
  left_join(bmi_results %>% select(pid, bmip, p50, p95, bmip95), by = "pid") %>%
  mutate(
    BMI_Score = case_when(
      bmip >= 5  & bmip < 85 ~ 100,
      bmip >= 85 & bmip < 95 ~ 70,
      bmip >= 95 & bmip95 < 120 ~ 30,
      bmip95 >= 120 & bmip95 < 140 ~ 15,
      bmip95 >= 140 ~ 0,
      TRUE ~ NA_real_
    )
  )

```

# Fasting glucose score
TODO: There is one individual with > 125 mg/dL

-   Non-diabetic with values \<100 mg/dL = `100`

-   Non-diabetic with values 100 to \<125 mg/dL = `60`

```{r}
summary(score_df$ResultGlucose)

score_df <- score_df %>%
  mutate(
    Fasting_Glucose_Score = case_when(
      ResultGlucose < 100 ~ 100,
      ResultGlucose >= 100 & ResultGlucose < 125 ~ 60,
      ResultGlucose >= 125 ~ 0,
      TRUE ~ NA_real_
    ),
    Fasting_Glucose_Flag = case_when(
      ResultGlucose >= 125 ~ "ERROR: glucose >=125",
      TRUE ~ NA_character_
    )
  )

```

# Non-HDL cholesterol score

-   \<100 mg/dL = `100`

-   100 to \<120 mg/dL = `60`

-   120 to \<145 mg/dL = `40`

-   145 to \<190 mg/dL = `20`

-   ≥190 mg/dL = `0`

```{r}
summary(score_df$ResultNonHDLCholesterol)

score_df <- score_df %>%
  mutate(
    NonHDL_Cholesterol_Score = case_when(
      ResultNonHDLCholesterol < 100 ~ 100,
      ResultNonHDLCholesterol >= 100 & ResultNonHDLCholesterol < 120 ~ 60,
      ResultNonHDLCholesterol >= 120 & ResultNonHDLCholesterol < 145 ~ 40,
      ResultNonHDLCholesterol >= 145 & ResultNonHDLCholesterol < 190 ~ 20,
      ResultNonHDLCholesterol >= 190 ~ 0,
      TRUE ~ NA_real_
    )
  )
```

# SBP (bpdia) and DBP (bpdia) score

-   \<90th percentile of age, sex, and height = `100`

-   SBP and DBP 90th to \<95th percentile = `75`

-   95th percentile to 95th percentile + 12 mmHg = `50`

-   95th percentile + 12 mmHg to \<95th percentile + 30 mmHg = `25`

-   ≥95th percentile + 30 mmHg = `0`

```{r}
library(pedbp)

pedbp_df <- score_df %>%
  mutate(
    male = case_when(
      sex == "male" ~ 1,
      sex == "female" ~ 0,
      TRUE ~ NA_real_
    )
  ) %>%
  dplyr::select(pid,
                childID,
                age_months,
                male,
                height_cm = height_cm_HSIIVisit,
                X_hs2_pdf_bpsys,
                X_hs2_pdf_bpdia) %>%
  filter(
    !is.na(X_hs2_pdf_bpsys),
    !is.na(X_hs2_pdf_bpdia),
    !is.na(male),
    !is.na(age_months),
    !is.na(height_cm)
  )

summary(pedbp_df$X_hs2_pdf_bpsys)  # SBP / systolic
summary(pedbp_df$X_hs2_pdf_bpdia)  # DBP / diastolic
summary(pedbp_df$age_months)  # DBP / diastolic
summary(pedbp_df$height_cm)  # DBP / diastolic


# bp_pct_results <- p_bp(
#   q_sbp = pedbp_df$X_hs2_pdf_bpsys,
#   q_dbp = pedbp_df$X_hs2_pdf_bpdia,
#   male   = pedbp_df$male,
#   age    = pedbp_df$age_months,
#   height = pedbp_df$height_cm,
#   source = "martin2022"
# )

# score_df <- pedbp_df %>%
#   mutate(
#     sbp_p = bp_pct_results$sbp_p,
#     dbp_p = bp_pct_results$dbp_p,
#     SBP_percentile = sbp_p * 100,
#     DBP_percentile = dbp_p * 100
#   )


bp_90 <- q_bp(
  p_sbp = 0.90,
  p_dbp = 0.90,
  male   = pedbp_df$male,
  age    = pedbp_df$age_months,
  height = pedbp_df$height_cm,
  source = "martin2022"
)

bp_95 <- q_bp(
  p_sbp = 0.95,
  p_dbp = 0.95,
  male   = pedbp_df$male,
  age    = pedbp_df$age_months,
  height = pedbp_df$height_cm,
  source = "martin2022"
)

score_df <- pedbp_df %>%
  mutate(
    SBP_90 = bp_90$sbp,
    DBP_90 = bp_90$dbp,
    SBP_95 = bp_95$sbp,
    DBP_95 = bp_95$dbp
  )

score_df <- score_df %>%
  mutate(
    Blood_Pressure_Score = case_when(
      is.na(X_hs2_pdf_bpsys) | is.na(X_hs2_pdf_bpdia) |
        is.na(SBP_90) | is.na(DBP_90) |
        is.na(SBP_95) | is.na(DBP_95) ~ NA_real_,
      
      X_hs2_pdf_bpsys >= SBP_95 + 30 |
        X_hs2_pdf_bpdia >= DBP_95 + 30 ~ 0,
      
      X_hs2_pdf_bpsys >= SBP_95 + 12 |
        X_hs2_pdf_bpdia >= DBP_95 + 12 ~ 25,
      
      X_hs2_pdf_bpsys >= SBP_95 |
        X_hs2_pdf_bpdia >= DBP_95 ~ 50,
      
      X_hs2_pdf_bpsys >= SBP_90 |
        X_hs2_pdf_bpdia >= DBP_90 ~ 75,
      
      X_hs2_pdf_bpsys < SBP_90 &
        X_hs2_pdf_bpdia < DBP_90 ~ 100,
      
      TRUE ~ NA_real_
    )
  )


```
